Robust Bayesian Analysis of Causal Inference Problems

### 2024-03-21 JASA-template.tex
Highlight potential issues with derivation of the treatment effect.

\subsection{Regression Model}
Regression methods are widely used in causal effect estimation. The
main idea behind these regression methods is to remove the
correlation between the treatment indicator and the error term
\cite{winship99,HECKMAN1985}. To do so, we rely on $p$ different observed quantities
or predictors denoted by $X\coloneqq$ $[X_1^T, \dots, X_n^T]^T$
where each $X_i\in\mathbb{R}^p$.
Each $X_i$ is treated as a $p$-dimensional row vector,
so $X$ is a $n\times p$ matrix.
Now, let
$\beta \coloneqq (\beta_1$, \dots, $\beta_p)^T$ denote the vector of regression
coefficients
related to the predictors, and $\beta_T$ denote a regression coefficient related to the treatment.
Then we can define a linear model for the outcome
so that
\begin{equation}
	Y_i =  T_i \beta_{T} + X_i\beta + \epsilon_i
\end{equation}
where $\epsilon_i\sim \mathcal{N}(0, \sigma^2)$. Clearly, when
the underlying true outcome model is linear with respect to the treatment,
%%% MT to BT: should we say $\delta_{n+1}$ here (and Y_{n+1}, T_{n+1})?
\begin{equation}
	\delta_i = \mathbb{E}(Y_i\mid T_i =1) - \mathbb{E}(Y_i\mid T_i=0) = \beta_{T}.
\end{equation}
To estimate $\beta_T$,
especially in the presence of confounders,
we also need to consider the
association between the treatment indicators and the predictors.
In literature, authors often suggest a probit link function to 
construct the regression model.
In this way, we can
specify the conditional probability that subject $i$ receives the treatment through a linear model. 
That is, for another vector of regression coefficients 
$\gamma\coloneqq(\gamma_1, \cdots, \gamma_p)^T$ we define
\begin{align}
	P(T_i=1\mid X_i) = \Phi(X_i\gamma)
\end{align}
where $\Phi(\cdot)$ denotes the cumulative distribution function
of a standard normal distribution. To incorporate this probit
link function, we assume that we can model the $T_i$ through the following \cite{albert93}:
\begin{align}
    T_i^* &= X_i\gamma +u_i \\
    T_i   &= \mathbb{I}(T_i^*>0)
    =
    \begin{cases}
    1 & \text{if }T_i^*>0 \\
    0 & \text{otherwise}
    \end{cases}
\end{align}
where $u_i\sim\mathcal{N}(0,1)$.

>>>>

\subsection{Regression Model}
Regression methods are widely used in causal effect estimation. The
main idea behind these regression methods is to remove the
correlation between the treatment indicator and the error term
\cite{winship99,HECKMAN1985}. To do so, we rely on $p$ different observed quantities
or predictors denoted by $X\coloneqq$ $[X_1^T, \dots, X_n^T]^T$
where each $X_i\in\mathbb{R}^p$.
Each $X_i$ is treated as a $p$-dimensional row vector,
so $X$ is a $n\times p$ matrix.
Now, let
$\beta \coloneqq (\beta_1$, \dots, $\beta_p)^T$ denote the vector of regression
coefficients
related to the predictors, and $\beta_T$ denote a regression coefficient related to the treatment.
Then we can define a linear model for the outcome
so that
\begin{equation}
	Y_i =  T_i \beta_{T} + X_i\beta + \epsilon_i
\end{equation}
where $\epsilon_i\sim \mathcal{N}(0, \sigma^2)$. Clearly, when
the underlying true outcome model is linear with respect to the treatment,
%%% MT to BT: should we say $\delta_{n+1}$ here (and Y_{n+1}, T_{n+1})?
\begin{equation}
	\delta_i = \mathbb{E}(Y_i\mid T_i =1) - \mathbb{E}(Y_i\mid T_i=0) = \beta_{T}.
\end{equation}
*** MT to TB: trying to derive this\dots ***      
\begin{align}
  \delta_i
  &=\mathbb{E}(Y_i\mid T_i =1) - \mathbb{E}(Y_i\mid T_i=0)\\
  &=\beta_T+\mathbb{E}(X_i\beta+\epsilon_i\mid T_i =1) - \mathbb{E}(X_i\beta+\epsilon_i\mid T_i=0)\\
  %%% MT to BT:
  \intertext{*** I get why $\epsilon_i$ has expectation zero (assumed independent from $T_i$); but why do the $X_i\beta$ terms vanish, since $X_i$ and $T_i$ are not assumed independent? maybe this can only assumed to hold for $i=n+1$ i.e. for the `next' person, not for the person in the data\dots ***}
  &=\beta_{T}.
\end{align}
To estimate $\beta_T$ from the data $X$, $Y$ and $T$,
especially in the presence of confounders,
we also need to consider the
association between the treatment indicators $T$ and the predictors $X$.
In literature, authors often suggest a probit link function to 
construct the regression model.
In this way, we can
specify the conditional probability that subject $i$ receives the treatment through a linear model. 
That is, for another vector of regression coefficients 
$\gamma\coloneqq(\gamma_1, \cdots, \gamma_p)^T$ we define
\begin{align}
	P(T_i=1\mid X_i) = \Phi(X_i\gamma)
\end{align}
where $\Phi(\cdot)$ denotes the cumulative distribution function
of a standard normal distribution. To incorporate this probit
link function, we assume that we can model the $T_i$ through the following \cite{albert93}:
\begin{align}
    T_i^* &= X_i\gamma +u_i \\
    T_i   &= \mathbb{I}(T_i^*>0)
    =
    \begin{cases}
    1 & \text{if }T_i^*>0 \\
    0 & \text{otherwise}
    \end{cases}
\end{align}
where $u_i\sim\mathcal{N}(0,1)$.
